---
title: "man"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# RVaggregator
Raster and Vector spatial aggregation tool

## Summary

RVaggregator aims to ease aggregation of spatial data. Aggregation involves the coarsening of a high resolution dataset (raster) across a lower resolution raster or polygon(s).

I have not found an "all in one" and relatively fast aggregation solution in e.g. R/python/gdal etc that I like. RVaggregator allows you to:
- Define a flexible range of statistics (cf. e.g. gdal which has a finite number of pre-compiled functions - it  only had "sum" introduced in ~2019)
- Parallelise processing across a range of input datasets
- Output neat shapefiles and 2D/3D rasters.
- Work with ordinal (e.g. land cover class) and continuous (e.g. heights) inputs
- Avoid the headache of having to make a specific bit of code for a specific aggregation

Also comes with command line interface wrapper in `CLI/RVaggregator-CLI.R`.

## Install

wip

## Command line use

wip

## R use

wip

``` {r}
library(RVaggregator)
library(terra)
```

``` {r, fig.height = 4, fig.width = 9, message = FALSE}
par(mfrow = c(1, 3))
plot(rast("data/sample/sample_input_raster_ordinal.tif"), main = "input_file raster with ordinal data")

plot(vect("data/sample/sample_shapefile/sample_shapefile.shp"), main = "Input aggregation_file polygon data")

plot(rast("data/sample/sample_input_raster_ordinal.tif"), main = "Both")
plot(vect("data/sample/sample_shapefile/sample_shapefile.shp"), add = TRUE)

aggregated_data <- RVaggregator(input_file = "data/sample/sample_input_raster_ordinal.tif",
             aggregation_file = "C:/Users/micromet/Desktop/locs.tif", #"data/sample/sample_raster_aggregate.tif",#"data/sample/sample_shapefile/sample_shapefile.shp",
             aggregation_type = "fraction",
             output_directory = "data/sample/output",
             poly_chunk_size = 5000)
```

The output data has same format as `aggregation_file`
```{r, message = FALSE}
class(aggregated_data)
```

Unlike `aggregation_file` , the polygons (each with unique ID) contain the data
```{r, message = FALSE}
head(aggregated_data)

```
