---
title: "RVaggregator"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# RVaggregator
Raster and Vector spatial aggregation tool

## Summary

RVaggregator aims to ease aggregation of spatial data. Aggregation involves the coarsening of a high spatial resolution raster across rasters or polygons of lower resolution.

I have not found an "all in one" and relatively fast aggregation solution in e.g. R/python/gdal etc that I like. RVaggregator can:

* Define a flexible range of aggregation statistics (cf. e.g. gdal which has a finite number of pre-compiled functions - it  only had "sum" introduced in ~2019)
* Output neat shapefiles and 2D/3D rasters.
* Work with ordinal (e.g. land cover class) and continuous (e.g. heights) inputs
* Aggregate data through command line (cf. R IDE or program)

Also comes with command line interface wrapper in `CLI/RVaggregator-CLI.R`.

## Install

Start up R and run:

``` {r, eval = FALSE}
install.packages("devtools")
devtools::install_github("willmorrison1/RVaggregator")
```

The required packages should then be installed.

## Command line use
The repo comes with a CLI wrapper e.g.

```
 R --no-save <RVaggregator-CLI.R
```

When run without arguments it will give a prompt on required inputs

```
usage: RVaggregator [--] [--help] [--opts OPTS] [--cache_directory CACHE_DIRECTORY] [--memory_fraction
       MEMORY_FRACTION] [--aggregation_chunk_size AGGREGATION_CHUNK_SIZE] [--aggregate_ordinal
       AGGREGATE_ORDINAL] input_file aggregation_file output_directory
```

Follow the R use to get the same variable names and descriptions required for the CLI inputs. 

## R use

``` {r, message = FALSE, warning = FALSE}
library(RVaggregator)
```

``` {r, fig.height = 4, fig.width = 9, message = FALSE, warning = FALSE}
par(mfrow = c(1, 3))
plot(rast("data/sample/input/sample_input_raster_ordinal.tif"), main = "input_file raster with ordinal data")

plot(vect("data/sample/input/sample_shapefile/sample_shapefile.shp"), main = "Input aggregation_file polygon data")

plot(rast("data/sample/input/sample_input_raster_ordinal.tif"), main = "Both")
plot(vect("data/sample/input/sample_shapefile/sample_shapefile.shp"), add = TRUE)

aggregated_data <- RVaggregator(input_file = "data/sample/input/sample_input_raster_ordinal.tif",
                                aggregation_file = "data/sample/input/sample_shapefile/sample_shapefile.shp",
                                aggregation_type = "fraction",
                                output_directory = "data/sample/output",
                                poly_chunk_size = 500)

```

The output data has same format as `aggregation_file`
```{r, message = FALSE, warning = FALSE}
class(aggregated_data)
```

Unlike `aggregation_file` , the polygons (each with unique ID) contain the data
```{r, message = FALSE, warning = FALSE}
head(aggregated_data)

```

You can also do the aggreagation using rasters, where `aggregation_file` raster has lower spatial resolution than
`input_file`.

``` {r, fig.height = 4, fig.width = 7, message = FALSE, warning = FALSE}
par(mfrow = c(1, 2))
plot(rast("data/sample/input/sample_input_raster_ordinal.tif"), main = "input_file raster\n with ordinal data")

plot(rast("data/sample/input/aggregation_file.tif"), main = "Input aggregation_file\nraster data")

aggregated_data <- RVaggregator(input_file = "data/sample/input/sample_input_raster_ordinal.tif",
                                aggregation_file = "data/sample/input/aggregation_file.tif",
                                aggregation_type = "fraction",
                                output_directory = "data/sample/output",
                                poly_chunk_size = 500)

```

When `aggregation_file` is raster, output is also raster with 3rd dimension with names of stats
``` {r}
class(aggregated_data)
names(aggregated_data)
```

Plot the second variable of the `aggregated_data`
``` {r, fig.height = 4, fig.width = 4, message = FALSE}
print(names(aggregated_data)[2])
plot(aggregated_data[[2]], main = "Aggregated data,\nsecond stat")

```
